scikit-image, a Python library for 2D and 3D image processing

### 2016-08-15 skimage_tomo.tex
New version with more files (figures and scripts to generate them)

\definecolor{listinggray}{gray}{0.1}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{
	backgroundcolor=\color{lbcolor},
	tabsize=4,
	rulecolor=,
	language=Python,
        basicstyle=\scriptsize,
        %upquote=true,
        aboveskip={1.5\baselineskip},
        columns=fixed,
        showstringspaces=false,
        extendedchars=true,
        breaklines=true,
        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
        frame=single,
        %showtabs=false,
        %showspaces=false,
        %showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

The acquisition time of synchrotron tomography images has decreased
dramatically over the last decade, from hours to
seconds~\cite{Maire2014}. New modalities such as single-bunch
imaging~\cite{Rack2014} provide a time resolution down to the nanosecond
for radiography. However, the time subsequently spent in processing the
images has not decreased as much, so that the outcome of a successful
synchrotron imaging campaign often takes weeks or even months to be
transformed into scientific results.

Transforming billions of pixels and voxels to a few meaningful figures
represents a tremendous data reduction. Often, the sequence of operations
needed to produce these data is not known beforehand, or might be altered
due to artifacts~\cite{Marone2010}, or to an unforeseen evolution of
the sample. Image processing necessarily involves trial and error phases
to choose the processing workflow. Therefore, image processing tools need
to offer at the same time enough flexibility of use, a variety of
algorithms, and efficient implementations to allow for fast iterations
while adjusting the workflow.

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ~\cite{Abramoff2004, Schneider2012}
and its distribution Fiji~\cite{Schindelin2012} is a popular tool for 2-D
and 3-D images, thanks to its intuitive menus and graphical tools, and
the wealth of plugins contributed by a vivid
community~\cite{Schindelin2015}. For 3-D images, commercial tools such as
Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- hence
limiting the programming task and the risk of bugs. Matlab (TM) and its image
processing toolbox are popular in the academic community of computer
vision and image processing. The Python language is widely used in the
scientific world and in synchrotron facilitiies. As a general-purpose
language, Python is used in synchrotrons for controlling device
servers~\cite{pytango}, reconstructing tomography volumes from
radiographies~\cite{Gursoy2014, Mirone2014}, and various data processing
packages for macromolecular cristallography~\cite{Adams2010} or
fluorescence analysis~\cite{pymca}.

\texttt{scikit-image}~\cite{Vanderwalt2014} is a general-purpose image
processing library for the Python language, and a component of the ecosystem of
Python scientific modules commonly known as Scientific
Python~\cite{Oliphant2007}. Like the rest of the ecosystem,
\texttt{scikit-image} is released under a permissive open-source license and is
available free of charge. \texttt{scikit-image} is compatible with both 2-D and
3-D images, so that it can be used for a large number of imaging modalities,
such as microscopy, radiography or tomography. In this article, we explain how
\texttt{scikit-image} can be used for processing data acquired in synchrotron
imaging experiments, with a strong focus on microtomography 3-D images. This
article does not intend to be a pedagogical tutorial on \texttt{scikit-image}
for X-ray imaging, but rather to explain the rationale behind the package, and
what can be done with \texttt{scikit-image}.

@@

Image are manipulated as numerical arrays of uniform
data type, a format that guarantees a fast access to the computer memory. The
n-dimensional (2-D, 3-D, \dots) numerical array object is provided by the
\texttt{NumPy} module~\cite{Vanderwalt2011}.

@@

The benefits of \texttt{scikit-image} for image processing come not only
from the features of the package alone, but also from the rich
environment of Scientific Python~\cite{Oliphant2007, Perez2011}.
Fig.~\ref{fig:ecosystem} illustrates how several components of this
ecosystem are used for a typical image processing workflow.

@@

While several interpreters are available to execute interactively Python
instructions and scripts, the most popular one in the scientific world is
the IPython interpreter~\cite{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates for example a debugger,
introspection and profiling methods, and online help. Several IDEs
(Integrated Development Environment) integrate the IPython interpreter,
together with other components such as a text editor. For example, Spyder 
is a popular IDE (see Fig.~\ref{fig:spyder}).

@@

The Jupyter notebook~\cite{Kluyver2016} is a web application which has
grown from the IPython project, although the two projects have now
diverged. Jupyter notebooks provide an interactive development
environment within an Internet browser, where live code can be enriched
by explanatory text, equations and visualizations (figure or not??).

\paragraph{NumPy arrays}

- advanced operations in a compact form, at minimal memory consumption

- slicing to reduce memory cost

- nD : 2D, 3D, ...

\paragraph{Visualization libraries}

\paragraph{Advanced toolkits for signal processing and data science}

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn}~\cite{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}, it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

@@

\begin{figure}
    \centerline{\includegraphics[width=0.99\columnwidth]{mayavi_porous}}
\caption{\csentence{Sample figure title.}
      A short description of the figure content
      should go here.}
\end{figure}

@@

The success of a software package aimed at end-users is strongly
correlated to the quality of its documentation (find citation).
\texttt{scikit-image} users have access to different kinds of
documentation. Functions of the API (define API here or above) are all
documented using the same standard~\cite{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data type, together with
explainations on what the function does and how to use it. Functions
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

@@

In addition, a popular documentation tool is the graphical gallery of
examples (url), a part of which is displayed in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by displaying graphical examples of image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). Thumbnails are links to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.

@@

Finally, tutorials on \texttt{scikit-image} are available in various places,
either as YouTube videos, or in the SciPy Lecture Notes~\cite{scipylecturenotes}, a comprehensive online book of Scientific Python tutorials.

@@

\texttt{scikit-image} is developed by a diverse team of volunteers, with
more than 170 developers that have contributed to the package. The
development process takes place on the GitHub platform
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, or report bugs or ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are proprosed to users on a regular basis.

@@

\section*{Competing interests}
  The authors declare that they have no competing interests.

@@

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{refs}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
\nocite{label}

>>>>

\definecolor{listinggray}{gray}{0.1}
\definecolor{lbcolor}{rgb}{0.9,0.92,0.95}
\lstset{
	backgroundcolor=\color{lbcolor},
	tabsize=4,
	rulecolor=,
	language=Python,
        basicstyle=\scriptsize,
        %upquote=true,
        aboveskip={1.5\baselineskip},
        columns=fixed,
        showstringspaces=false,
        extendedchars=true,
        breaklines=true,
        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
        frame=single,
        %showtabs=false,
        %showspaces=false,
        %showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

@@

The acquisition time of synchrotron tomography images has decreased
dramatically over the last decade, from hours to
seconds~\citep{Maire2014}. New modalities such as single-bunch
imaging~\citep{Rack2014} provide a time resolution down to the nanosecond
for radiography. However, the time subsequently spent in processing the
images has not decreased as much, so that the outcome of a successful
synchrotron imaging campaign often takes weeks or even months to be
transformed into scientific results.

Transforming billions of pixels and voxels to a few meaningful figures
represents a tremendous data reduction. Often, the sequence of operations
needed to produce these data is not known beforehand, or might be altered
due to artifacts~\citep{Marone2010}, or to an unforeseen evolution of
the sample. Image processing necessarily involves trial and error phases
to choose the processing workflow. Therefore, image processing tools need
to offer at the same time enough flexibility of use, a variety of
algorithms, and efficient implementations to allow for fast iterations
while adjusting the workflow.

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ~\citep{Abramoff2004, Schneider2012}
and its distribution Fiji~\citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community~\citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- hence
limiting the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilitiies. As a
general-purpose language, Python is used in synchrotrons for controlling
device servers~\citep{pytango}, reconstructing tomography volumes from
radiographies~\citep{Gursoy2014, Mirone2014}, and various data
processing packages for macromolecular cristallography~\citep{Adams2010},
azimuthal integration of diffraction data~\citep{Ashiotis2015} or
fluorescence analysis~\citep{pymca}.

\texttt{scikit-image}~\citep{Vanderwalt2014} is a general-purpose image
processing library for the Python language, and a component of the ecosystem of
Python scientific modules commonly known as Scientific
Python~\citep{Oliphant2007}. Like the rest of the ecosystem,
\texttt{scikit-image} is released under a permissive open-source license and is
available free of charge. \texttt{scikit-image} is compatible with both 2-D and
3-D images, so that it can be used for a large number of imaging modalities,
such as microscopy, radiography or tomography. In this article, we explain how
\texttt{scikit-image} can be used for processing data acquired in synchrotron
imaging experiments, with a strong focus on microtomography 3-D images. This
article does not intend to be a pedagogical tutorial on \texttt{scikit-image}
for X-ray imaging, but rather to explain the rationale behind the package, and
what can be done with \texttt{scikit-image}.

@@

Image are manipulated as numerical arrays of uniform
data type, a format that guarantees a fast access to the computer memory. The
n-dimensional (2-D, 3-D, \dots) numerical array object is provided by the
\texttt{NumPy} module~\citep{Vanderwalt2011}.

@@

The benefits of \texttt{scikit-image} for image processing come not only
from the features of the package alone, but also from the rich
environment of Scientific Python~\citep{Oliphant2007, Perez2011}.
Fig.~\ref{fig:ecosystem} illustrates how several components of this
ecosystem are used for a typical image processing workflow.

@@

While several interpreters are available to execute interactively Python
instructions and scripts, the most popular one in the scientific world is
the IPython interpreter~\citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates for example a debugger,
introspection and profiling methods, and online help. Several IDEs
(Integrated Development Environment) integrate the IPython interpreter,
together with other components such as a text editor. For example, Spyder 
is a popular IDE (see Fig.~\ref{fig:spyder}).

@@

The Jupyter notebook~\citep{Kluyver2016} is a web application which has
grown from the IPython project, although the two projects have now
diverged. Jupyter notebooks provide an interactive development
environment within an Internet browser, where live code can be enriched
by explanatory text, equations and visualizations (figure or not??).

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all one-liners NumPy code. To
illustrate the compactness of NumPy code, modifying pixels values below a
threshold can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
using the possibility to index arrays with boolean arrays, called
\emph{masking}. Also, \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, which is a important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array:
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}

\paragraph{Visualization libraries.}

Visualizing images is an important component of the image processing
workflow, in order to verify the final result and to adjust the
parameters of intermediate processing operations.
\texttt{matplotlib}~\citep{Hunter2007} is the most popular 2D plotting
library of the Python ecosystem. It can be used to visualize 2D data such
as color or grayscale images, and 1D data such as contour lines, outlines
of segmented regions, histograms of gray levels, etc. Although
\texttt{matplotlib} has simple 3D plotting capabilities, it is
recommended to use the \texttt{mayavi} module~\citep{Ramachandran2011}
for applications requiring advanced 3D visualization, such as tomography. 
\texttt{mayavi} is based on the VTK toolkit. It exposes a simple API for
visualizing data passed as \texttt{numpy} arrays. For more advanced
visualizations, a large majority of VTK capabilities can be accessed
thanks to a pipeline API. A visualization pipeline can be designed either
using the pipeline API, or graphically in the pipeline window.
\texttt{mayavi} offers a good trade-off between simplicity of use for
common operations, and more advanced possibilities, including responsive
visualizations.

speak about skimage's visualizer as well?

\paragraph{Advanced toolkits for signal processing and data science.}

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn}~\citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}, it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{tomo_gallery}}
\caption{\csentence{Sample figure title.}
      A short description of the figure content
      should go here.}
\end{figure*}

@@

The success of a software package aimed at end-users is strongly
correlated to the quality of its documentation (find citation).
\texttt{scikit-image} users have access to different kinds of
documentation. Functions of the API (define API here or above) are all
documented using the same standard~\citep{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data type, together with
explainations on what the function does and how to use it. Functions
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

@@

In addition, a popular documentation tool is the graphical gallery of
examples (\url{http://scikit-image.org/docs/dev/auto_examples/}), a part of which is displayed in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by showcasing graphical examples of image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). Thumbnails are links to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.

@@

Finally, tutorials on \texttt{scikit-image} are available in various
places, either as YouTube videos, or in the SciPy Lecture
Notes~\citep{scipylecturenotes}, a comprehensive online book of Scientific
Python tutorials.

@@

\texttt{scikit-image} is developed by a diverse team of volunteers, with
more than 170 developers that have contributed to the package. The large
number of developers and users is a key element of the sustainability of
\texttt{scikit-image}. The
development process takes place on the GitHub platform
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, or report bugs or ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are proposed to users on a regular basis.

\section*{Conclusion}

@@

\section*{Competing interests}
: The authors declare that the research was conducted
in the absence of any commercial or financial relationships that could be
construed as a potential conflict of interest.

@@

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{refs}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}
